Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-10-21 Paper/ms.tex
Writing

\section{Orbit Simulations}
The expected (noise-free) signal due to an exoplanet is periodic, but
non-sinusoidal when the orbit is not perfectly circular. The expected
shape of the variations can be obtained from mechanical considerations.
To save time, we pre-computed the properties of orbits as a function of
ellipticity.

Consider a test particle moving in the $x$-$y$ plane under the influence of a
point mass at the origin. The equations of motion for the particle are:
\begin{eqnarray}
\frac{d^2x}{dt^2} &=& -\frac{x}{r^3} \\
\frac{d^2y}{dt^2} &=& -\frac{y}{r^3} \\
\end{eqnarray}
where $r = \sqrt{x^2 + y^2}$. The ODEs can be solved for any given set of
initial conditions for $x$, $y$, $\frac{dx}{dt}$, and $\frac{dy}{dt}$ to yield
the shape of the orbit. To simulate orbits
of different ellipticities, we set the initial position to $(1, 0)$, and
the initial velocity to $(0, v)$ where $v \in [0.4, 1]$.
If $v=1$, the orbit is circular and as $v$ decreases the orbit becomes more
elliptical. For trial values of $v$ ranging from 0.4 to 1 in steps of 0.005,
we simulated an orbit using a leapfrog integrator, and saved the
velocities $\frac{dx}{dt}$ and $\frac{dy}{dt}$ as a function of time to disk.
These saved orbits were used as a lookup table for constructing the expected
signal $y(t)$ due to a single planet.
Because of the initial conditions used,
the simulated orbits were all horizontally
aligned.

>>>>

\section{Orbit Simulations}
The expected (noise-free) signal due to an exoplanet is periodic, but
non-sinusoidal when the orbit is not perfectly circular. The expected
shape $m(t)$ of the variations can be obtained from mechanical considerations.
To save time, we pre-computed the properties of orbits as a function of
ellipticity.

Consider a test particle moving in the $x$-$y$ plane under the influence of a
point mass at the origin. The motion of the test particle represents the
reflex motion of the host star orbiting around the center of mass of the
system. The equations of motion for the particle are:
\begin{eqnarray}
\frac{d^2x}{dt^2} &=& -\frac{x}{r^3} \\
\frac{d^2y}{dt^2} &=& -\frac{y}{r^3} \\
\end{eqnarray}
where $r = \sqrt{x^2 + y^2}$. These ODEs can be solved for any given set of
initial conditions for $x$, $y$, $\frac{dx}{dt}$, and $\frac{dy}{dt}$ to yield
the shape of the orbit. To simulate orbits
of different ellipticities, we set the initial position to $(1, 0)$, and
the initial velocity to $(0, v)$ where $v \in [0.4, 1]$.
If $v=1$, the orbit is circular and as $v$ decreases the orbit becomes more
elliptical. For trial values of $v$ ranging from 0.4 to 1 in steps of 0.005,
we simulated an orbit using a leapfrog integrator, and saved the
velocities $\frac{dx}{dt}$ and $\frac{dy}{dt}$ as a function of time to disk.
These saved orbits were used as a lookup table for constructing the expected
signal $y(t)$ due to a single planet.
Because of the initial conditions, the simulated orbits were all horizontally
aligned. If the observer is located on the $x$-axis a large distance
from the origin, they will measure $m(t) = \dot{x}(t)$. However, if the
observer is located at

Although we parameterised the orbits using the initial velocity $v$,
the standard eccentricity $\epsilon$ can be recovered using the relation
$\epsilon = 1 - v^2$.
